---
title: "Homework 3"
author: "Philip Kim"
date: "10/12/2021"
output: 
  github_document:
    toc: true
    toc_depth: 6
---

```{r, message = FALSE}
library(tidyverse)
library(knitr)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

# Problem 1

```{r}
data("instacart")
```

## Dataset description

The instacart dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

## How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisle_df = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))
```

There are 134 aisles and the aisles with the most items ordered from are fresh vegetables and fresh fruits aisles with 150,609 and 150,473 orders, respectively.

## Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
aisle_df %>% 
  filter(n_obs > 10000) %>% 
  mutate(
    aisle = as.factor(aisle),
    aisle = fct_reorder(aisle, (n_obs))) %>% 
  ggplot((aes(x = aisle, y = n_obs))) + 
  labs(
    title = "Item Popularity in Instacart",
    x = "Aisle Name",
    y = "Number of Ordered Items") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_col() +
  coord_flip()
```


## Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(
    aisle == "baking ingredients" | 
    aisle == "dog food care" | 
    aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_obs = n()) %>% 
  mutate(rank = rank(-n_obs)) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  kable(caption = "Most popular items in Baking Ingredients, Dog Food Care, and Packaged Vegetables Fruits")
```


## Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
  filter(product_name == "Pink Lady Apples" |
         product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  mutate(
    order_dow = recode(order_dow, "0" = "Sun",
                                  "1" = "Mon",
                                  "2" = "Tues",
                                  "3" = "Wed",
                                  "4" = "Thurs",
                                  "5" = "Fri",
                                  "6" = "Sat"),
    order_dow = as.factor(order_dow)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) %>% 
  kable(caption = "Mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week")
```


# Problem 2

```{r}
data("brfss_smart2010")
```

## Data Cleaning

```{r}
brfss_tidy_smart2010 =
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health",
         response == "Excellent" |
         response == "Very good" |
         response == "Good" |
         response == "Fair" |   
         response == "Poor") %>% 
  mutate(response = as.factor(response),
         response = ordered(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

## In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_tidy_smart2010 %>% 
  filter(year == 2002 |
         year == 2010) %>%
  group_by(year, locationabbr) %>% 
  summarize(n_obs = n_distinct(locationdesc)) %>% 
  filter(n_obs >= 7) %>% 
  arrange(year, (n_obs)) %>% 
  kable()
```

There were a total of 6 states that were observed at 7 or more locations in 2002 which include CT, FL, MA, NC, NJ, and PA. Relatively, there were a total of 14 states that were observed at 7 or more locations in 2010 which include CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA.

## Spaghetti Plot

```{r}
brfss_tidy_smart2010 %>% 
  filter(response == "Excellent") %>% 
  group_by(year, locationabbr) %>% 
  summarize(mean_data_value = mean(data_value, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_data_value)) +
  labs(
    x = "Year",
    y = "Average Data Value within State",
    title = "Average Data Value of those with \nExcellent Overall Health over Time by State") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group = locationabbr))
```

## Two-Panel Plot

```{r}
brfss_tidy_smart2010 %>% 
  filter(year == 2006 |
         year == 2010,
         locationabbr == "NY") %>% 
  group_by(year, response) %>% 
  summarize(data_value = data_value) %>% 
  ggplot(aes(x = response, y = data_value)) +
  labs(
    x = "Response",
    y = "Data Value",
    title = "Distribution of Data Value by Response Type in 2006 and 2010") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_boxplot() +
  facet_grid(. ~ year)
```


# Problem 3

## Load and Tidy Data

```{r}
accel_df = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute_of_day",
    names_prefix = "activity_",
    values_to = "activity_count") %>% 
  mutate(
    weekend = if_else(day == "Saturday" | day == "Sunday", "weekend", "weekday"),
    minute_of_day = as.numeric(minute_of_day),
    day = as.factor(day),
    day = ordered(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))) %>% 
  relocate(week, day_id, weekend, everything())
```

.....


## Total Activity per Day

```{r}
accel_df %>% 
  group_by(week, day) %>% 
  summarize(total_activity_count = sum(activity_count)) %>% 
  kable()
```

....


## 24-hour activity time course plot

```{r}
accel_df %>% 
  mutate(
    hour = minute_of_day %/% 60) %>% 
  group_by(day_id, day, hour) %>% 
  summarize(activity_count = sum(activity_count)) %>% 
  ggplot(aes(x = hour, y = activity_count, color = day)) +
  labs(
    x = "Hour of the Day",
    y = "Activity Count",
    title = "24 Hour Activity Time Course per Day",
    color = "Day of the Week") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group = day_id))
```

....